---
title: "entropy_kde2d()"
output:
  rmarkdown::html_vignette:
    fig_width: 5
    fig_height: 4
    self_contained: false
vignette: |
  %\VignetteIndexEntry{entropy_kde2d()} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc}
bibliography: references.bib
csl: apa.csl
---

```{r echo = FALSE, warning=FALSE}
library(YEAB)
```

## Introduction

The `entropy_kde2d()` function estimates the Shannon entropy for a two-dimensional dataset using kernel density estimation (KDE). This function provides a non-parametric measure of entropy, useful for analyzing the uncertainty or randomness in bivariate distributions, such as spatial data of animal trajectories [e.g, @maei2009].

The parameters for `entropy_kde2d()` are as follows:

- `x`: A numeric vector representing the random variable in the first dimension.

- `y`: A numeric vector representing the random variable in the second dimension.

- `n_grid`: A numeric value specifying the number of grid cells used to evaluate the density; defaults to 150. Higher values result in finer density grids but increase computational complexity.

The function outputs a single numeric value representing the entropy of the given data.

## Example

```{r}
set.seed(123)
# Generate a 2D normal distribution with a correlation of 0.6
n <- 1000
mean <- c(0, 0)
sd_x <- 1
sd_y <- 5
correlation <- 0.6
sigma <- matrix(
  c(
    sd_x^2,
    correlation * sd_x * sd_y,
    correlation * sd_x * sd_y,
    sd_y^2
  ),
  ncol = 2
)
library(MASS)
simulated_data <- mvrnorm(n, mu = mean, Sigma = sigma)
x <- simulated_data[, 1]
y <- simulated_data[, 2]
# Plot the data
plot(simulated_data)
# Compute entropy using normal entropy formula
cov_matr <- cov(cbind(x, y))
sigmas <- diag(cov_matr)
det_sig <- prod(sigmas)
```

`normal_entropy` is a function that computes the entropy of a bivariate normal distribution given the number of dimensions `k`, the value of $\pi$, and the determinant of the covariance matrix `det_sig`. This is used, for example, in @maei2009 to compute the entropy of mice trajectories in a Morris water maze.

```{r}
normal_entropy <- function(k, pi, det_sig) {
  (k / 2) * (1 + log(2 * pi)) + (1 / 2) * log(det_sig)
}

entropia <- normal_entropy(k = 2, pi = pi, det_sig)
print(entropia) # Expected value close to 4.3997

# Compute entropy using entropy_kde2d
result <- entropy_kde2d(x, y, n_grid = 50)
print(result) # Expected value close to 4.2177
```

## References
