---
title: "event_extractor()"
output:
  rmarkdown::html_vignette:
    fig_width: 5
    fig_height: 4
    self_contained: false
vignette: |
  %\VignetteIndexEntry{event_extractor()} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc}
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(YEAB)
```

## Introduction

The proprietary data output format from the MED-PC software suite for experimental chambers is often coded in a `time.event` manner, in which users assign arbitrary numeric codes for each type of event, such that when analyzing the data there are specific events marking the start and end of each individual trail.

Here we present a function to slice MED-PC data based on start and end events. This function should be used after `read_med()`, which outputs a data frame or/and `.csv` (depending on how it's used) with 2 columns: time and events (in that order). The function returns a data frame with *n* x 4 columns: 

- `time`: the time of the event.
- `events`: the event numeric ID.
- `cum_id`: the index of each extracted trail (note that these will be in a *i·2+1* sequence due how the internal processing is done).
- `evname`: the assigned identifier for the extracted trails.

The `event_extractor()` function include the following parameters:

- `data_df`: a data frame of *n* rows x 2 columns, where columns corresponds to time and events IDs, in that order.
- `ev0`: start event ID (where the trials we want to extract begins).
- `ev1`: end event ID (where the trials we want to extract ends).
- `evname`: a string for the event name, for identification purposes.

## Example

First let's load a raw MED-PC file from an experimental session and assign it to a data frame:

```{r tidy = FALSE, message = FALSE, warning = FALSE}
data("fi60_raw_from_med")
temp_file <- tempfile(fileext = ".txt")
writeLines(fi60_raw_from_med, temp_file)

example_processed <- read_med(
  fname = temp_file, save_file = FALSE,
  path_save = path_to_save, col_r = "C:", out = TRUE,
  col_names = c("time", "event"), num_col = 6, time_dot_event = TRUE
)

head(example_processed, 10)
```

In this specific example session we have four types of trials: 

- FI15 with start code *5* and end code *3*
- FI30 with start code *55* and end code *33*
- Peak 15s with start code *4* and end code *41*
- Peak 30s with start code *44* and end code *42*

If we wanted to extract the FI15 trials, we would do it like this:

```{r}
extracted_FI15 <- event_extractor(
  data_df = example_processed,
  ev0 = 5, ev1 = 3,
  evname = "IF15"
)

head(extracted_FI15, 30)
```

Now let's assume we want to extract all four types of trails and store them in separated data frames. We can do that using a `for` cycle like this:

```{r}
extracted_trials_df <- list() # A list to store the resulting data frames.

param_list <- list( # A list of parameters for each trial type.
  list(ev0 = 5, ev1 = 3, evname = "IF15"),
  list(ev0 = 55, ev1 = 33, evname = "IF30")
)

for (parameters in param_list) {
  trial_df <- event_extractor(
    data_df = example_processed,
    ev0 = parameters$ev0,
    ev1 = parameters$ev1,
    evname = parameters$evname
  )
  extracted_trials_df[[length(extracted_trials_df) + 1]] <- trial_df # Store the extracted data frame in the data frames list.
}
```

Here's a preview of how each data frame would look like:

```{r echo = FALSE}
for (i in seq_along(extracted_trials_df)) {
  cat("Data frame", i, ":\n")
  print(head(extracted_trials_df[[i]], 5))
}
```

After this process you can analyze each trial data separately in whichever way you need.
