---
title: "sample_from_density()"
output:
  rmarkdown::html_vignette:
    fig_width: 5
    fig_height: 4
    self_contained: false
vignette: |
  %\VignetteIndexEntry{sample_from_density()} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc}
---

```{r echo = FALSE, warning=FALSE}
library(YEAB)
```

## Introduction

Behavioral data often exhibit complex patterns that may not adhere to well-known parametric distributions. Traditional statistical methods reliant on such distributions may therefore prove inadequate in capturing the nuanced behaviors observed in real-world scenarios.

Approximating a density distribution from a data sample using Kernel Density Estimation allows researchers to estimate the underlying patterns that could emerge from a specific experiment, thus, the ability to generate synthetic data samples that closely mimic the distribution of observed behavioral data serves as a valuable tool in behavioral analysis. Such samples allow researchers to explore various hypotheses, validate statistical models, and assess the efficacy of experimental interventions in a controlled setting.

Here we introduce the `sample_from_density()` function which generates a given `n` amount of data points from a density distribution calculated using KDE (Kernel Density Estimation) from a given data set.

The function takes two parameters:

- `x` A numeric vector of data points from an distribution.
- `n` the number of samples to return.

## Example

Let's generate a random sample from a normal distribution:

```{r}
set.seed(142)
normal_sample <- rnorm(100)
```

Now let's create a sample of 100 data points from the distribution estimated with the `sample_from_density()` function:

```{r}
sampled_dist <- sample_from_density(normal_sample, 100)
```

Finally let's compare both distributions:
```{r, echo = FALSE}
plot(ks::kde(normal_sample),
  xlab = "", main = "original vs sample_from_density()",
  col = "blue",
  ylim = c(0, 0.5)
)
kde_sample <- ks::kde(sampled_dist)
plot(kde_sample, col = "red", add = TRUE)
legend("topleft",
  c("sample_from_density", "original"),
  lty = c(1, 1), col = c("red", "blue")
)
```
