---
title: "unit_normalization()"
output:
  rmarkdown::html_vignette:
    fig_width: 5
    fig_height: 4
    self_contained: false
vignette: |
  %\VignetteIndexEntry{unit_normalization()} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc}
---

```{r echo = FALSE, warning=FALSE}
library(YEAB)
```

## Introduction

When analyzing numeric data, either discrete or continuous variables, it is often necessary or at least practical to normalize the values in order to get a more comprehensible scale to analyze the data in, this is, transforming the values to a $0 ≤ x ≤ 1$ scale, where $0$ is the lowest value and $1$ the highest in the distribution. 

We included two functions to normalize and rescale numeric vectors, `unit_normalization()` and `ab_range_normalization()`, respectively. The former takes a numeric vector `x` as input and outputs a normalized version of the same distribution. 

## Example

```{r}
x <- 5:45
x_scaled <- unit_normalization(x)
x_scaled
```

Similarly the `ab_range_normalization()` function can be used to rescale a numeric vector `x` to an arbitrary range between `a` and `b`. E.g.:

```{r}
x <- 5:45
a <- 1
b <- 100
x_scaled <- ab_range_normalization(x, a, b)
x_scaled
```

```{r}
x <- rnorm(1000)
hist(x, main = "Original Distribution")
x_scaled <- unit_normalization(x)
hist(x_scaled, main = "Normalized Distribution")
```
